Sentiment classification of online political discussions: a comparison of a word-based and dependency-based method

### 2014-04-23 manuscript.tex
Added old version of dependence path section

Over the past years online political discussions have received a lot of attention. E.g. the Obama 2012 election team initiated an extensive use of text analytics and machine learning techniques towards online material to guide advertising campaigns, identifying key voters, and improve fundraising \cite{Issenberg12}. There has also been a lot of concern about the alarming growth in hate and racism against minorities like Muslims, Jews and Gypsies in online discussions \cite{r6,s2}. Sentiment analysis (SA) is the discipline of automatically determining sentiment in text material and may be one important tool in understanding the diversity of opinions on the Internet. SA has mostly been used to analyze opinions in comments and reviews about commercial products, but there are also examples of SA towards political tweets and discussions, see e.g. Tumasjan et al. \shortcite{Tumasjan2010}; Chen et al. \shortcite{Chen10}. SA of political discussions is known to be a difficult task since citations, irony and sarcasm is very common \cite{Bing12}.

@@

@@

In this section we present two methods to classify sentences as either positive, neutral or negative towards a target word. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Both target words, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters are words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. Let $tw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the target word in the sentence. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ represent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(tw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function \textbf{imp} computes the importance of the sentiment word $sw_{j}$ on the target word appearance $tw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters are within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the sentiment towards the target word to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set, as described in section \ref{sec:exp} below.

@@

>>>>

Over the past years online political discussions have received a lot of attention. E.g. the Obama 2012 election team initiated an extensive use of text analytics and machine learning techniques towards online material to guide advertising campaigns, identifying key voters, and improve fundraising \cite{Issenberg12}. There has also been a lot of concern about the alarming growth in hate and racism against minorities like Muslims, Jews and Gypsies in online discussions \cite{r6,s2}. Sentiment analysis (SA) is the discipline of automatically determining sentiment in text material and may be one important tool in understanding the diversity of opinions on the Internet.

@@

\section{Previous work}

Sentiment classification aims to classify a document or sentence as either positive or negative and sometimes also neutral. There are mainly two approaches, one based on machine learning and one based on based on using a list of words with given sentiment scores (lexicon-based approach). For machine learning any existing method can be used, e.g. na\"{\i}ve Bayes and support vector machine, \cite{Joachims99,Shawe00}. One simple lexicon-based approach is to count the number of words with positive and negative sentiment in the document as suggested by Hu and Liu \shortcite{Hu04}. One may classify the opinion of larger documents like movie or product reviews or smaller documents like tweets, comments or sentences. See Liu \shortcite{Bing12}, chapters three to five and references therein for the description of several opinion classification methods.

SA has mostly been used to analyze opinions in comments and reviews about commercial products, but there are also examples of SA towards political tweets and discussions, see e.g. Tumasjan et al. \shortcite{Tumasjan2010}; Chen et al. \shortcite{Chen10}. SA of political discussions is known to be a difficult task since citations, irony and sarcasm is very common \cite{Bing12}.

@@

In this section we present two methods to classify sentences as either positive, neutral or negative towards a target word. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Both target words, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters are words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. Let $tw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the target word in the sentence. Note that we only consider one target word at the time. E.g. if a sentence contains two target words, e.g. Quran and Islam, the sentence is first classified with respect to Quran and then with respect to Islam. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ represent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(tw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function \textbf{imp} computes the importance of the sentiment word $sw_{j}$ on the target word appearance $tw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters are within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the sentiment towards the target word to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set, as described in section \ref{sec:exp} below.

@@

---------------------------------------------------

A second way of determining the importance of a sentiment word towards a target based on syntactically parsed texts, is to establish a list of grammatical dependency paths between words, and test whether such paths exist between the target and the sentiment words in the material under investigation \cite{Jiang11}. The assumption would be that, if there is a meaningful grammatical relation between a target and a sentiment word, it is likely that they are semantically related to each other. Furthermore, it is reasonable to expect than some paths are stronger indicators of the overall sentiment of the sentence than others. To test this method, we have made a list of 42 grammatical dependency paths and given them a score between 1-3. The higher the score is, the better indicator of sentiment the path is assumed to be. In the following paragraphs, we will present our reasoning behind the choice of paths and the weight we have given them. The paths are written as follows: postag-target:postag-sentiment word{\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn({\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn etc.). \emph{Up} and \emph{dn} show whether you move up or down in the dependency tree. So the relation between a noun target and a verb sentiment word where the noun is the subject of the verb, is represented as subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up. The relation between a target noun and a sentiment noun where the former is the subject and the latter is the direct object of the same verb, is subst:subst{\textunderscore}{\textunderscore}SUBST{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn.

	A first group consists of paths from subject targets to sentiment predicates. Such paths can e.g. go from a subject to a verbal predicate, subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up, as in (\ref{verbpred}), or from a subject to an adjectival or nominal predicate in the context of a copular verb, as in (\ref{adpred}), subst:adj{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn or subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn.

	\begin{examples}
	\item\label{verbpred}
	\gll Ja, n{\aa} \textbf{jubler(sent.w.)} vel \textbf{muslimene(target)}.
	yes now exult probably muslims+the
	\glt 'Yes, the muslims probably exult now.'
	\glend

	\item\label{adpred}
	\gll [...] for meg er \textbf{Muhammed(target)} like \textbf{levende(sent.w.)} som min far [...].
	{} for me is Muhammad as alive as my father {}
	\glt 'To me, Muhammad is as much alive as my dad is.'
	\glend

	\end{examples}

	We have chosen to give subject-predicate paths the highest possible score, 3. Firstly, the combination of a subject and a predicate will result in a proposition, a statement which is evaluated as true or false \cite{Heim98}. We expect that a proposition typically will represent the opinion of the speaker, although e.g. irony and certain kinds of embedding can shift the truth evaluation in some cases \cite{Hooper73}. Secondly, if the predicate represents an event brought about by an intentional agent, such as \emph{jubler}, 'exult', in (\ref{verbpred}), the subject will typically represent that agent. If the predicate has a positive or negative sentiment, we expect that this sentiment is directed towards this intentional agent.

	A second group we have considered, contains paths from subject targets to sentiment words embedded within the predicate. Examples of such paths are those from the subject to the nominal direct object of a verb, as in (\ref{subjobj}), subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn. Paths from subjects into different kinds of adverbials are also a part of this group.

	\begin{examples}
	\item\label{subjobj}
	\gll [...] ekstreme \textbf{muslimer(target)} beg{\aa}r \textbf{voldshandlinger(sent.w.)} [...].
	{} extreme muslims commit {violent acts} {}
	\glt 'Extreme muslims commit violent acts.'
	\glend
	\end{examples}

	We consider paths from subjects to objects good indicators of sentiment and therefore give them the highest score, 3. The reasoning is much the same as for subject predicate paths: The statement is a proposition and the subject will often be the agent of the event. Also, the object, being an obligatory argument of the verb, is presumably closely semantically connected to it. Paths from subjects to adverbials, on the other hand, get a lower score, 2. This is because our experience is that the parser performs less well on adverbials. Also, adverbials are not obligatory arguments, and we therefore expect more semantic variation than for objects.

	The paths in our third group go from targets to sentiment words within the predicate. These include paths from nominal direct object target to verbal predicates, as in (\ref{objverb}), subst:verb{\textunderscore}{\textunderscore}DOBJ{\textunderscore}up, and from various kinds of adverbials to verbal predicates, among other things:

	\begin{examples}
	\item\label{objverb}
	\gll En Norsk dommer skal kontant \textbf{avvise(sent.w.)} \textbf{Sharia(target)} [...].
	a Norwegian judge shall strictly reject sharia {}
	\glt 'A Norwegian judge must strictly reject sharia.'
	\glend
	\end{examples}

	We assume that predicate-internal paths are less good indicators of sentiment than the above groups, as predicates need to combine with a subject to form a proposition. Also, arguments within the predicate usually do not represent the intentional agent of the event. Such paths will get the score 1.

	Our fourth and final group of dependency paths contains paths internal to the nominal phrase, such as from target nouns to attributive adjectives, as in (\ref{adjatr}), subst:adj{\textunderscore}{\textunderscore}ATR{\textunderscore}dn, and from target complements of attributive prepositions to target nouns, as in (\ref{substatr}), subst:subst{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}up:

	\begin{examples}
	\item\label{adjatr}
	\gll [...] \textbf{fundamentalistisk(sent.w.}) \textbf{islam(target)} [...]
	{} fundamentalist Islam {}
	\glt 'fundamentalist Islam'
	\glend

	\item\label{substatr}
	\gll [...] \textbf{kampen(sent.w} mot \textbf{islam(target)} [...]
	{} fight+the against Islam {}
	\glt 'the fight against Islam'
	\glend

	\end{examples}

	We suspect paths internal to the nominal phrase to be relatively good indicators of sentiment: A positively or negatively qualified noun will probably often represent the sentiment of the speaker. At the same time, a nominal phrase of this kind can be used in many different contexts where the holder of the sentiment is not the speaker, so we therefore expect such paths to be somewhat less good indicators than subject-predicate paths. We therefore assign them the score 2.

	Some complex paths receive a lower score than those indicated here, to compensate for parser inaccuracies.

---------------------------------------------------
